Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2025-01-17 article.tex
Updates to Cooks membrane

Figure \ref{fig:mms_disp_accuracy}(a) shows the displacement magnitude discretisation errors ($L_2$ and $L_\infty$) as a function of the average cell width for the three mesh types (hexahedral, tetrahedral and polyhedral), while Figure \ref{fig:mms_disp_accuracy}(b) shows the corresponding order of accuracy plots.
For ease of interpretation, the symbol shapes in the figures have been chosen to correspond to the cell shapes: a square for hexahedra, a triangle for tetrahedra and a pentagon for polyhedra.
The maximum ($L_\infty$) and average ($L_2$) displacement discretisation errors are seen to reduce at an approximately second-order rate for all mesh types, except for the $L_2$ error on the hexahedral meshes, which is approximately 2.3 on the finest grid.
%where, interestingly, the errors are, on average, smaller on the polyhedral meshes.
%The reason for this is unclear but may be due to the stencil of the structured polyhedra being more isotropic and containing close neighbours.
\begin{figure}[htbp]
	\centering
	\subfigure[Displacement magnitude discretisation errors]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_dispErrors_v2} 
   	}
	\subfigure[Displacement discretisation error order of accuracy]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_disp_orderOfAccuracy_v2}  
   	}
	\caption{Manufactured solution cube case: the accuracy and order of accuracy for displacement magnitude on the \emph{regular} meshes}
	\label{fig:mms_disp_accuracy}
\end{figure}

@@

\paragraph{Case 4: Cook's membrane}
Cook's membrane (Figure \ref{fig:cooks_membrane}) is a well-known bending-dominated benchmark case used in linear and non-linear analysis.
The 2-D plane strain tapered panel (trapezoid) is fixed on one side and subjected to uniform shear traction on the opposite side.
The vertices of the trapezoid (in mm) are (0, 0), (48, 44), (48, 60),  and (0, 44).
The problem is solved quasi-statically, using 30 equally-sized loading increments, and there are no body forces.
%\begin{figure}[htbp]
%   \centering
%   \includegraphics[width=0.4\textwidth]{figures/cooksMembrane-geometry} 
%   \caption{Cook's membrane case geometry (dimensions in mm) and loading conditions (fixed left, traction $\tau$ applied to right)}
%   \label{fig:cooks_membrane}
%\end{figure}
\begin{figure}[htbp]
	\centering
	\subfigure[Case geometry and dimensions (in mm)]
	{
		\label{fig:cooks_membrane_geometry}
   		\includegraphics[scale=1]{figures/cooksMembrane-geometry} 
   	}
   	\qquad
	\subfigure[Quadrilateral mesh with 144 cells]
	{
		\label{fig:cooks_membrane_mesh}
   		\includegraphics[scale=0.14]{figures/cooksMembrane-mesh}  
   	}
	\caption{Cook's membrane case geometry and mesh}
   \label{fig:cooks_membrane}
\end{figure}

The current work considers three forms of the problem:
\begin{enumerate}[label=\roman*.]
	\item Small strain, linear elastic \cite{Zienkiewicz2000, Simplas}: $E=70$ MPa, $\nu=1/3$, and $\tau = 6250$ Pa.
	\item Finite strain, neo-Hookean hyperelastic \cite{Pelteret2018}: $E=1.0985$ MPa, $\nu=0.3$, and $\tau = 0.0625$ MPa.
	\item Finite strain, neo-Hookean hyperelastoplastic \citep{Simo1992, Simplas, Cesar2001}: $E=206.9$ MPa, $\nu=0.29$, $\sigma_y = 0.45 + 0.12924\bar{\varepsilon}_p + (0.715 - 0.45)(1- e^{-16.93\bar{\varepsilon}_p})$ MPa, and $\tau = 0.3125$ MPa.
\end{enumerate}
where $E$ is the Young's modulus, $\nu$ is the Poisson's ratio, $\sigma_y$ is the yield strength, $\bar{\varepsilon}_p$ is the equivalent plastic strain, and $\tau$ is the prescribed shear traction.
%The Young's modulus $E = 206.9$ MPa and Poisson's ratio $\nu=0.29$, with the yield stress $\sigma_y$ given as \citep{Simo1992}
%\begin{eqnarray}
%	\sigma_y = \sigma_Y + H\bar{\varepsilon}_p + (\sigma_{\infty} - \sigma_Y)(1- e^{-\delta\bar{\varepsilon}_p})
%\end{eqnarray}
%The plastic yielding parameters are $\sigma_Y = 0.45$ MPa, $\sigma_{\infty} = 0.715$ MPa, $\delta = 16.93$, and $H = 0.12924$ MPa, where the hardening variable $\bar{\varepsilon}_p$ corresponds to equivalent plastic strain.
%\noindent
%\textbf{Sources for case i:} \cite{Zienkiewicz2000}\\
%O.C. Zienkiewicz, R.L. Taylor. The finite element method. Butterworth Heinemann, 2000.\\
%https://www.simplassoftware.com/benchmarks.html\#biblio-58\\
%https://github.com/spolanski/CoFEA/tree/master/benchmarks/02-cooks-membrane\\
%\textbf{Sources for case ii:} \cite{dealII95}\\
%https://www.dealii.org/current/doxygen/deal.II/code\_gallery\_Quasi\_static\_Finite\_strain\_Compressible\_Elasticity.html\\
%\textbf{Sources for case iii:} \citep{Simo1992}
Eight successively refined quadrilateral meshes are considered, where the cell counts are 9, 36, 144 (Figure \ref{fig:cooks_membrane}(b)), 576, $2\,304$, $9\,216$, $36\,864$, and $147\,456$.
%mesh.1 - 3x3 - 9 CVs\\
%mesh.2 - 6x6 - 36 CVs\\
%mesh.3 - 12x12 - 144 CVs\\
%mesh.4 - 24x24 - 576 CVs\\
%mesh.5 - 48-48 - 2304 CVs\\
%mesh.6 - 96x96 - 9216 CVs\\
%mesh.7 - 192x192 - 36864 CVs
%mesh.8 - 384x384 - 147456 CVs

Figure \ref{fig:cooksMembrane_sigmaEq} shows the predicted equivalent stress distribution on the mesh with $36\,842$ cells for linear elastic, hyperelastic and hyperelastoplastic cases.
The stress distribution is consistent with bending, with regions of high stress near the upper and lower surfaces and a line of relatively unstressed material in the centre.
The greatest equivalent stresses occur at the top-left corner and the lower surface in all versions of the case.
In the hyperelastoplastic case, almost the entire domain is plastically yielding with only a small thin region remaining elastic, indicated by the blue line in Figure \ref{fig:cooksMembrane_sigmaEq}(c).
\begin{figure}[htbp]
   \centering
	\subfigure[Linear elastic case \hl{kPa}]
	{
		\includegraphics[width=0.3\textwidth]{figures/cooksMembrane-hookean-sigmaEq.pdf}  
	}
	\subfigure[Hyperelastic case]
	{
		\includegraphics[width=0.3\textwidth]{figures/cooksMembrane-neoHookean-sigmaEq.pdf}  
	}
	\subfigure[Hyperelastoplastic case]
	{
		\includegraphics[width=0.3\textwidth]{figures/cooksMembrane-neoHookeanPlastic-sigmaEq.pdf}  
	}
   \caption{Equivalent (von Mises) stress distribution for the three Cook's membrane cases using the mesh with $36\,842$ cells}
   \label{fig:cooksMembrane_sigmaEq}
\end{figure}

@@

The deformed configuration of the beam at six time steps is shown in Figure \ref{fig:dynamic_cantilever}(a) for the mesh with $2\,160$ cells, where the upper surface is seen to drop below the horizontal from approximately t = 0.2 s to 0.32 s.
The corresponding displacement magnitude of the centre of the upper surface of the beam vs time is shown in Figure \ref{fig:dynamic_cantilever}(b), where the predictions are seen to converge to the results from finite element software Abaqus (element code C3D8) using the $138\,240$ mesh.
%Although the current case does not have a reference solution, it has been chosen to test the robustness of the solution procedures when large rotations and strains are present.
\begin{figure}[htbp]
   \centering
%   \includegraphics[width=0.5\textwidth]{figures/cantilever} 
	\subfigure[\hl{update}Displacement magnitude for t = $\left\{0, \, 0.1, \, 0.2\right\}$ s (translucent) and t = $\left\{0.3, \, 0.4, \, 0.5 \right\}$ s (opaque) for the mesh with $138\,240$ cells]
	{
	   \includegraphics[width=0.5\textwidth]{figures/cantilever_deformed_v2}
   	} \quad
	\subfigure[Displacement magnitude of the centre of the upper surface of the beam vs time]
	{
	   \includegraphics[width=0.6\textwidth]{figures/cantilever_disp} 
   	   %\includegraphics[height=0.45\textwidth]{figures/cantilever_sigmaEq}  
   	}
   \caption{Deflection of a 3-D Cantilevered Beam}
   \label{fig:dynamic_cantilever}
\end{figure}
%\hl{It would be good to add a reference results, e.g. from FE}

>>>>

Figure \ref{fig:mms_disp_accuracy}(a) shows the displacement magnitude discretisation errors ($L_2$ and $L_\infty$) as a function of the average cell width for the three mesh types (hexahedral, tetrahedral and polyhedral), while Figure \ref{fig:mms_disp_accuracy}(b) shows the corresponding order of accuracy plots.
For ease of interpretation, the symbol shapes in the figures have been chosen to correspond to the cell shapes: a square for hexahedra, a triangle for tetrahedra and a pentagon for polyhedra.
The maximum ($L_\infty$) and average ($L_2$) displacement discretisation errors are seen to reduce at an approximately second-order rate for all mesh types, except for the $L_2$ error on the hexahedral meshes, which is approximately 2.3 on the finest grid.
%where, interestingly, the errors are, on average, smaller on the polyhedral meshes.
%The reason for this is unclear but may be due to the stencil of the structured polyhedra being more isotropic and containing close neighbours.
\begin{figure}[htbp]
	\centering
	\subfigure[Displacement magnitude discretisation errors]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_dispErrors_v2} 
   	}
	\subfigure[Displacement discretisation error order of accuracy]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_disp_orderOfAccuracy_v2}  
   	}
	\caption{Manufactured solution cube case: the accuracy and order of accuracy for displacement magnitude}
	\label{fig:mms_disp_accuracy}
\end{figure}

@@

\paragraph{Case 4: Cook's membrane}
Cook's membrane (Figure \ref{fig:cooks_membrane}) is a well-known bending-dominated benchmark case used in linear and non-linear analysis.
The 2-D plane strain tapered panel (trapezoid) is fixed on one side and subjected to uniform shear traction on the opposite side.
The vertices of the trapezoid (in mm) are (0, 0), (48, 44), (48, 60),  and (0, 44).
%\begin{figure}[htbp]
%   \centering
%   \includegraphics[width=0.4\textwidth]{figures/cooksMembrane-geometry} 
%   \caption{Cook's membrane case geometry (dimensions in mm) and loading conditions (fixed left, traction $\tau$ applied to right)}
%   \label{fig:cooks_membrane}
%\end{figure}
\begin{figure}[htbp]
	\centering
	\subfigure[Case geometry and dimensions (in mm)]
	{
		\label{fig:cooks_membrane_geometry}
   		\includegraphics[scale=1]{figures/cooksMembrane-geometry} 
   	}
   	\qquad
	\subfigure[Quadrilateral mesh with 144 cells]
	{
		\label{fig:cooks_membrane_mesh}
   		\includegraphics[scale=0.14]{figures/cooksMembrane-mesh}  
   	}
	\caption{Cook's membrane case geometry and mesh}
   \label{fig:cooks_membrane}
\end{figure}

The current work considers three forms of the problem:
\begin{enumerate}[label=\roman*.]
	\item Small strain, linear elastic \cite{Zienkiewicz2000, Simplas}: $E=70$ MPa, $\nu=1/3$, and $\tau = 6.25$ kPa.
	\item Finite strain, neo-Hookean hyperelastic \cite{Pelteret2018}: $E=1.0985$ MPa, $\nu=0.3$, and $\tau = 62.5$ kPa.
	\item Finite strain, neo-Hookean hyperelastoplastic \citep{Simo1992, Simplas, Cesar2001}: $E=206.9$ MPa, $\nu=0.29$, $\sigma_y = 0.45 + 0.12924\bar{\varepsilon}_p + (0.715 - 0.45)(1- e^{-16.93\bar{\varepsilon}_p})$ MPa, and $\tau = 312.5$ kPa.
\end{enumerate}
where $E$ is the Young's modulus, $\nu$ is the Poisson's ratio, $\sigma_y$ is the yield strength, $\bar{\varepsilon}_p$ is the equivalent plastic strain, and $\tau$ is the prescribed shear traction.
%The Young's modulus $E = 206.9$ MPa and Poisson's ratio $\nu=0.29$, with the yield stress $\sigma_y$ given as \citep{Simo1992}
%\begin{eqnarray}
%	\sigma_y = \sigma_Y + H\bar{\varepsilon}_p + (\sigma_{\infty} - \sigma_Y)(1- e^{-\delta\bar{\varepsilon}_p})
%\end{eqnarray}
%The plastic yielding parameters are $\sigma_Y = 0.45$ MPa, $\sigma_{\infty} = 0.715$ MPa, $\delta = 16.93$, and $H = 0.12924$ MPa, where the hardening variable $\bar{\varepsilon}_p$ corresponds to equivalent plastic strain.
%\noindent
%\textbf{Sources for case i:} \cite{Zienkiewicz2000}\\
%O.C. Zienkiewicz, R.L. Taylor. The finite element method. Butterworth Heinemann, 2000.\\
%https://www.simplassoftware.com/benchmarks.html\#biblio-58\\
%https://github.com/spolanski/CoFEA/tree/master/benchmarks/02-cooks-membrane\\
%\textbf{Sources for case ii:} \cite{dealII95}\\
%https://www.dealii.org/current/doxygen/deal.II/code\_gallery\_Quasi\_static\_Finite\_strain\_Compressible\_Elasticity.html\\
%\textbf{Sources for case iii:} \citep{Simo1992}
Eight successively refined quadrilateral meshes are considered, where the cell counts are 9, 36, 144 (Figure \ref{fig:cooks_membrane}(b)), 576, $2\,304$, $9\,216$, $36\,864$, and $147\,456$.
%mesh.1 - 3x3 - 9 CVs\\
%mesh.2 - 6x6 - 36 CVs\\
%mesh.3 - 12x12 - 144 CVs\\
%mesh.4 - 24x24 - 576 CVs\\
%mesh.5 - 48-48 - 2304 CVs\\
%mesh.6 - 96x96 - 9216 CVs\\
%mesh.7 - 192x192 - 36864 CVs
%mesh.8 - 384x384 - 147456 CVs
The problem is solved quasi-statically, and there are no body forces.
The linear elastic and hyperelastic cases use one loading increment, whereas the hyperelastoplastic case uses 30 equally-sized loading increments.

Figure \ref{fig:cooksMembrane_sigmaEq} shows the predicted equivalent stress distribution on the mesh with $36\,842$ cells for linear elastic, hyperelastic and hyperelastoplastic cases.
The stress distribution is consistent with bending, with regions of high stress near the upper and lower surfaces and a line of relatively unstressed material in the centre.
The greatest equivalent stresses occur at the top-left corner and the lower surface in all versions of the case.
In the hyperelastoplastic case, almost the entire domain is plastically yielding with only a small thin region remaining elastic, indicated by the blue line in Figure \ref{fig:cooksMembrane_sigmaEq}(c).
\begin{figure}[htbp]
   \centering
	\subfigure[Linear elastic case]
	{
		\includegraphics[width=0.28\textwidth]{figures/cooksMembrane-hookean-sigmaEq.pdf}  
	}
	\subfigure[Hyperelastic case]
	{
		\includegraphics[width=0.3\textwidth]{figures/cooksMembrane-neoHookean-sigmaEq.pdf}  
	}
	\subfigure[Hyperelastoplastic case]
	{
		\includegraphics[width=0.3\textwidth]{figures/cooksMembrane-neoHookeanPlastic-sigmaEq.pdf}  
	}
   \caption{Equivalent (von Mises) stress distribution for the three Cook's membrane cases using the mesh with $36\,842$ cells}
   \label{fig:cooksMembrane_sigmaEq}
\end{figure}

@@

The deformed configuration of the beam at six time steps is shown in Figure \ref{fig:dynamic_cantilever}(a) for the mesh with $2\,160$ cells, where the upper surface is seen to drop below the horizontal from approximately t = 0.2 s to 0.32 s.
The corresponding displacement magnitude of the centre of the upper surface of the beam vs time is shown in Figure \ref{fig:dynamic_cantilever}(b), where the predictions are seen to converge to the results from finite element software Abaqus (element code C3D8) using the $138\,240$ mesh.
%Although the current case does not have a reference solution, it has been chosen to test the robustness of the solution procedures when large rotations and strains are present.
\begin{figure}[htbp]
   \centering
%   \includegraphics[width=0.5\textwidth]{figures/cantilever} 
	\subfigure[Displacement magnitude for t = $\left\{0, \, 0.1, \, 0.2\right\}$ s (translucent) and t = $\left\{0.3, \, 0.4, \, 0.5 \right\}$ s (opaque) for the mesh with $138\,240$ cells]
	{
	   \includegraphics[width=0.5\textwidth]{figures/cantilever_deformed_v2}
   	} \quad
	\subfigure[Displacement magnitude of the centre of the upper surface of the beam vs time]
	{
	   \includegraphics[width=0.6\textwidth]{figures/cantilever_disp} 
   	   %\includegraphics[height=0.45\textwidth]{figures/cantilever_sigmaEq}  
   	}
   \caption{Deflection of a 3-D Cantilevered Beam}
   \label{fig:dynamic_cantilever}
\end{figure}
%\hl{It would be good to add a reference results, e.g. from FE}
